---
title: "CMU Reproducible Research Contest Submission"
author: "Ryan Elmore and Greg J. Matthews"
date: "9/25/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(xtable)
library(lme4)
```

## Data

We merged three data sets together in order to complete this analysis. 
Unfortunately, one source is proprietary and we can't share the three raw data
sets. However, we are able to share the final merged version. 

```{r data}
df <- readRDS("../data/bangs-merged-final.rds") %>%
  mutate(has_bangs = if_else(has_bangs == "y", "Yes", "No"))
```

## Table One

The following code produces Table 1 in Elmore and Matthews. 

```{r, results = "asis", message = F, warning = F, include = F}
tab_one <- df %>% 
  group_by(pi_pitch_group, has_bangs) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n))
print(xtable(tab_one, caption = "Table 1 from Elmore and Matthews."), type = "latex")
```

### Chi-Square Test

```{r}
chisq.test(table(df$has_bangs, df$pi_pitch_group))
```

## Table Two

```{r, results = "asis", message = F, warning = F}
df <- df %>% 
  dplyr::mutate(., I_swing = ifelse(swing == "swing", 1, 0))
prt_tab_one <- xtable(table(df$has_bangs, df$I_swing),
                      caption="Table 2 (counts) from Elmore and Matthews", digits = c(3, 3, 3))
print(prt_tab_one, type = "latex")
```

```{r, results = "asis", message = F, warning = F}
df <- df %>% 
  dplyr::mutate(., I_swing = ifelse(swing == "swing", 1, 0))
prt_tab_one <- xtable(prop.table(table(df$has_bangs, df$I_swing), 1),
                      caption="Table 2 (proportions) from Elmore and Matthews", digits = c(3, 3, 3))
print(prt_tab_one, type = "latex")
```

## Odds Ratios Between Tables 1 and 2

```{r}
questionr::odds.ratio(table(df$has_bangs, df$I_swing))
```

## Swing Model Results

Some more data manipulation:

```{r}
df <- readRDS("../data/bangs-merged-final.rds") %>%
  dplyr::rename(., description = description.x) %>% 
  dplyr::mutate(.,
                has_bangs = if_else(has_bangs == "y", "Yes", "No"),
                count = paste(ball, "-", strike, sep = ""),
                is_swing = ifelse(swing == "swing",1,0),
                is_miss = ifelse(call_code %in% c("S","W"), 1, 0),
                is_contact = ifelse(call_code %in% c("S","W"), 0, 1),
                is_foul = ifelse(description %in% 
                                   c("Foul", "Foul Tip", "Foul (Runner Going)"),1, 0),
                #is_fastball = ifelse(pitch_category == "FB", 1, 0), 
                is_fastball = ifelse(pi_pitch_group == "FA", 1, 0),
                batter_mlbid = as.character(batter_mlbid))
```

```{r, results = "asis"}
swing_model <- glmer(is_swing ~ is_fastball + has_bangs + count + 
                       cs_prob + (1|batter_mlbid),
                     family = binomial, data = df)
xtable(summary(swing_model)$coef, digits = rep(4, 5))
```

### Confidence Interval

```{r, cache = T}
confint(swing_model, parm = "has_bangsYes")
```

## Contact Model

```{r}
swings <- df %>% 
  dplyr::filter(., is_swing == 1)
dim(swings)[1]
```

### Table 3

```{r, results = "asis"}
contact_model <- glmer(is_contact ~  cs_prob + is_fastball*has_bangs + 
                         (1 + has_bangs|batter_mlbid)  + (1|mlbid), 
                       data = swings, 
                       family = "binomial")
xtable(summary(contact_model)$coef, digits = rep(4, 5))
```

### Confidence Intervals

```{r, cache = T}
exp(0.5905)
confint(contact_model, "has_bangsYes")

ids <- swings[!duplicated(swings$batter_mlbid),
              c("batter_mlbid", "batter_first", "batter_last")]
r_effects <- data.frame(batter_mlbid = 
                          row.names(coef(contact_model)$batter_mlbid), 
                        int = coef(contact_model)$batter_mlbid$`(Intercept)`,
                        slope = coef(contact_model)$batter_mlbid$has_bangsYes)
r_effects <- merge(ids, r_effects, by.x = "batter_mlbid", by.y = "batter_mlbid")
r_effects <- r_effects[order(-r_effects$slope), ]
r_effects$name <- paste(r_effects$batter_first, r_effects$batter_last)
```

```{r, results = "asis"}
print(xtable(r_effects[,c("name","int","slope")]), include.rownames=FALSE)
```

```{r, cache = T, results = "asis", warning = F, message = F, eval = F}
#Bootstrap to get CI for differences to start.  
set.seed(2017)
odds <- function(p){
  return(p/(1-p))
} 

OR_offspeed <- OR_fastball <- c()
ids <- swings[!duplicated(swings$batter_mlbid),
              c("batter_mlbid", "batter_first", "batter_last")]
#Bootstrap 
nsim <- 500

for (i in 1:nsim){print(i)
  ind <- sample(1:nrow(swings), nrow(swings), replace = TRUE)
  swings_boot <- swings[ind, ]
  
  contact_model_boot <- glmer(is_contact ~  cs_prob + is_fastball*has_bangs + 
                                (1 + has_bangs|batter_mlbid)  + (1|mlbid), 
                              data = swings_boot, 
                              family = "binomial")
  
  r_effects <- data.frame(batter_mlbid =
                            row.names(coef(contact_model_boot)$batter_mlbid),
                          int = coef(contact_model_boot)$batter_mlbid$`(Intercept)`,
                          slope = coef(contact_model_boot)$batter_mlbid$has_bangsYes)
  r_effects <- merge(ids, r_effects, by.x = "batter_mlbid", by.y = "batter_mlbid")
  r_effects <- r_effects[order(-r_effects$slope), ]
  r_effects$name <- paste(r_effects$batter_first, r_effects$batter_last)
  if (i == 1){
    int_boot <- merge(ids, r_effects[, c("batter_mlbid","int")], 
                      by.x = "batter_mlbid", 
                      by.y = "batter_mlbid" , 
                      all.x = TRUE)
  } else {
    int_boot <- merge(int_boot, r_effects[, c("batter_mlbid", "int")], 
                      by.x = "batter_mlbid", 
                      by.y = "batter_mlbid" , 
                      all.x = TRUE)
  }
  
  if (i == 1){
    slope_boot <- merge(ids, r_effects[, c("batter_mlbid", "slope")], 
                        by.x = "batter_mlbid", 
                        by.y = "batter_mlbid", 
                        all.x = TRUE)
  } else {
    slope_boot <- merge(slope_boot, r_effects[, c("batter_mlbid", "slope")], 
                        by.x = "batter_mlbid", 
                        by.y = "batter_mlbid" , 
                        all.x = TRUE)
  }
  
  newdat <- data.frame(cs_prob = median(swings$cs_prob), 
                       is_fastball = c(0,0,1,1), 
                       has_bangs = c("No","Yes","No","Yes"))
  mm <- model.matrix(~ cs_prob+ is_fastball*has_bangs , newdat) ## create
  newdat$y <- mm%*%fixef(contact_model_boot)
  newdat$p <- exp(newdat$y)/(1+exp(newdat$y))
  
#  predict(contact_model_boot, newdat, re.form = NA, type = "response") #would give the same results
  
  OR_offspeed[i] <- odds(newdat$p[newdat$is_fastball == 0 & 
                                    newdat$has_bangs == "Yes"]) /
    odds(newdat$p[newdat$is_fastball == 0 & newdat$has_bangs == "No"])
  
  OR_fastball[i] <- odds(newdat$p[newdat$is_fastball == 1 & 
                                    newdat$has_bangs == "Yes"]) /
    odds(newdat$p[newdat$is_fastball == 1 & newdat$has_bangs == "No"])
  }

quantile(OR_offspeed, c(0.025, 0.975))
quantile(OR_fastball, c(0.025, 0.975))

quantile(log(OR_offspeed), c(0.025, 0.975))


r_effects <- data.frame(batter_mlbid = row.names(coef(contact_model)$batter_mlbid), int = coef(contact_model)$batter_mlbid$`(Intercept)`,slope = coef(contact_model)$batter_mlbid$has_bangsYes)
r_effects <- merge(ids,r_effects, by.x = "batter_mlbid",by.y = "batter_mlbid")
r_effects <- r_effects[order(-r_effects$slope),]
r_effects$name <- paste(r_effects$batter_first, r_effects$batter_last)

ci_rand_slope <- cbind(slope_boot[,1:3],t(apply(slope_boot[,-c(1:3)],1,function(x){exp(quantile(x, c(0.025, 0.975)))})))

r_effects <- merge(r_effects,ci_rand_slope, by.x = "batter_mlbid", by.y = "batter_mlbid",all.x = TRUE)

r_effects$OR_out <- paste0(round(exp(r_effects$slope),3)," (",round((r_effects$`2.5%`),3),", ",round((r_effects$`97.5%`),3),")")
r_effects <- r_effects[order(-r_effects$slope),]

print(xtable(r_effects[,c("name","OR_out")]), include.rownames=FALSE)
```

## Exit Velocity Model

```{r, include = F}
contacts <- swings %>% 
  dplyr::filter(., is_contact == 1 )

speed_model <- lmer(launch_speed ~   cs_prob + is_fastball + has_bangs   + (1|batter_mlbid) + (1|mlbid)  , data = contacts)
summary(speed_model)
confint(speed_model)
coef(speed_model)
```

### Table 4

```{r, results = "asis"}
print(xtable(summary(speed_model)$coef, caption = "Table 4 in Elmore Matthews"))
```

### Table 4 P-Values

```{r}
2*(1 - pnorm(summary(speed_model)$coef[,3]))
```

## Session Information

```{r}
sessionInfo()
```

